Effect of character correlation

### 2017-09-19 Analysis/EffectCorrelationStats.Rmd
streamlined stats analysis

```{r, eval = TRUE, message = FALSE}
library(knitr)
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

@@

@@

```{r}
## Running the pairwise Wilcox tests:
sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
```

@@

```{r, echo = FALSE, eval = FALSE}
library(xtable)
## Wrapper function
make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {

  table <- xtable(table, digit = digit, caption = caption, label = label)

  if(longtable == TRUE) {
    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
  } else {
    print(table, include.rownames = FALSE)
  }
}

## Labels
labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

@@

```{r}
## getting the aov data
aov_data_RF_best <- get.aov.data(whole_data, "RF", "norm", combined = TRUE)
aov_data_Tr_best <- get.aov.data(whole_data, "Triples", "norm", combined = TRUE)
aov_data_RF_null <- get.aov.data(whole_data, "RF", "null", combined = TRUE)
aov_data_Tr_null <- get.aov.data(whole_data, "Triples", "null", combined = TRUE)

## Models with normalised tree score function of scenario (with nestedness)
model_RF_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_best)
model_Tr_best <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_best)
model_RF_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_RF_null)
model_Tr_null <- lmer(NTS ~ scenario + (1|method) + (1|taxa) + (1|character),
                      data = aov_data_Tr_null)

## Testing model assumptions
## Normality of the residuals
norm_RF_best <- shapiro.test(residuals(model_RF_best))
norm_Tr_best <- shapiro.test(residuals(model_Tr_best))
norm_RF_null <- shapiro.test(residuals(model_RF_null))
norm_Tr_null <- shapiro.test(residuals(model_Tr_null))
normality <- list(norm_RF_best, norm_Tr_best, norm_RF_null, norm_Tr_null)

## Variance homoscedasticity
homo_RF_best <- bartlett.test(NTS ~ scenario, data = aov_data_RF_best)
homo_Tr_best <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_best)
homo_RF_null <- bartlett.test(NTS ~ scenario, data = aov_data_RF_null)
homo_Tr_null <- bartlett.test(NTS ~ scenario, data = aov_data_Tr_null)
homoscedasticity <- list(homo_RF_best, homo_Tr_best, homo_RF_null, homo_Tr_null)

# ## Summarise the results
results <- t(rbind(sapply(normality, function(x) x[c(1,2)]), 
                   sapply(homoscedasticity, function(x) x[c(1,2,3)])))
results <- matrix(as.numeric(results), ncol = 5)
colnames(results) <- c("W", "p.value", "Bartlett's K^2", "df", "p.value")
rownames(results) <- c("RF best", "Triples best", "RF null", "Triples null")
results <- round(results, 3)

# ## Check assumption's validities
check.validity <- function(X) {ifelse(X[2] >= 0.05 && X[5] >= 0.05, TRUE, FALSE)}
validity <- apply(results, 1, check.validity)

## Report results
results <- cbind(as.data.frame(results), validity)
```

```{r, echo = FALSE}
kable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

@@

```{r, eval = FALSE, echo = FALSE}
xtable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

>>>>

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

@@

Effect of character correlation per scenario
-----------------

```{r, warning = FALSE}
## Pooling the data per scenarios
scenario_RF_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "RF", best = "norm")
scenario_Tr_norm <- multi.pool(whole_data, param = c("maxi", "mini", "rand"),
                               metric = "Tr", best = "norm")
scenario_RF_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
                               metric = "RF", best = "rand")
scenario_Tr_null <- multi.pool(whole_data, param = c("maxi", "mini", "norm"),
                               metric = "Tr", best = "rand")

## Preparing the table labels
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)

## Combining the data into a table
scenarios <- pool.table(list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null),
                    comp.label = comp.labels, metric.label = metric.labels, label.param = "scenario")
```

```{r, echo = FALSE}
library(knitr)
## Setting the table captions
captions_pooled_scenarios <- paste("Summary statistics of the normalised distances to the best or null tree for the pooled scenarios.")

## Printing the tables
kable(scenarios, row.names = FALSE, caption = captions_pooled_scenarios)
```

```{r, eval = FALSE, echo = FALSE}
library(xtable)
make.xtable(scenarios, caption = captions_pooled_scenarios, label = "Tab_pooledscenarios")
```

### Testing the effect of the scenario on the NTS

```{r}
## Applying an lm to all the data
results <- test.effect.aov(whole_data, formula = NTS ~ scenario, fun = lm)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

Applying a Bhattacharrya Coefficient and Wilcoxon test

```{r, warning = FALSE}
## Summarise into a table
scenario_list <- list(scenario_RF_norm, scenario_Tr_norm, scenario_RF_null, scenario_Tr_null)
comp.labels <- c("Best", rep("", 5), "Null", rep("", 5))
metric.labels <- rep(c("RF", "", "", "Tr", "", ""), 2)
test_table <- pair.test.table(scenario_list, comp.label = comp.labels, metric.label = metric.labels)
```

```{r, echo = FALSE}
## Caption
cap_pool_scen_test <- paste("Difference between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correciton)")

## Printing the tables
kable(test_table, row.names = FALSE, caption = cap_pool_scen_test)
```

```{r, eval = FALSE, echo = FALSE}
make.xtable(test_table, caption = cap_pool_scen_test, label = "Tab_pooledscenarios_test")
```

Effect of character correlation per number of character
-----------------

Effect of character correlation per number of taxa
-----------------

Effect of character correlation per methods
-----------------

@@

```{r}
## Summarising the data
sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
```

@@

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

@@

```{r}
## Applying an lm to all the data with error
results <- test.effect.aov(whole_data,
  formula = NTS ~ scenario + (1|method) + (1|taxa) + (1|character), fun = lmer)
```

Checking the model assumptions

```{r, echo = FALSE}
kable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

@@

```{r, eval = FALSE, echo = FALSE}
xtable(results$validity, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```
